Hybrid models as an alternative to traditional Convolutional Neural Networks for image classification

### 2022-11-25 doc/paper.tex
Start on stochastic pca, more text

Convolutional Neural Networks are neural networks which are specialized to work on images, or other types of data where the structure of the data carries meaning. A normal neural network treats each input indepentently, and inputs are flattened such that any spatial information is lost. CNNs, on the other hand, explicitly take the spatial dimensions into account, by using convolutional kernels and pooling layers as part of their architecture.

@@

A decision tree is an intiutive way of performing both classification and regression tasks, by asking a series of yes or no questions about an instance in the dataset. These question will split the original dataset, the root node, into child nodes which contain subsets of the dataset, and these nodes will be split again, creating a tree structure which terminates in leaf nodes which imply a choice of class.

In a binary classification problem, the algorithm for creating a decision tree is quite simple. First, consider each feature of the dataset, and attempt to split all instances into two child nodes based on a simple true/false statement about this feature. Having done this for every feature, calculate the degree of homogeneity\footnote{This can described by calculating the Gini-factor, or by calculating the entropy of the set} of the two child nodes created, when considering the actual class of the instances. The feature which created the most homegenous split, is the feature which has the largest effect on the final class, and as such should be the first feature we use to split the dataset. After this, we have two subsets, and we perform the same algorithm on these subsets again. This algorithm terminates when it is not possible to split the data further, because all nodes only contain instances belonging to the same class. In practice, it is advised to stop the algorithm earlier by introducing limits to the depth of the tree, or limits to how few instances can be in one node. Once the algorithm has terminated, we have a series of leaf nodes containing subsets of the data.

@@

To create a random forest, we perform a bootstrap of the original dataset, creating many similar instances with small permutations introduced by resampling with replacement. We now create one decision tree per bootstrap, only using a random subset of the total amount of features every time we perform a split\footnote{The number of features used is often set to $\sqrt{p}$}. The result of this is that we create many slightly different trees, each slightly suboptimal in their splits. These trees give predictions which are worse than simply creating a single decision tree using all the features and all the data, but when basing our prediction of the majority vote of all the trees, we are able to make predictions which are more generalizable and less prone to overfitting.

@@

The MNIST dataset is well known multiclass classification problem used to test machine learning algorithms. It consists of $70000$ black and white images of hand drawn digits, each $28$ by $28$ pixels. Each image is labeled with one of ten classes, corresponding to the digit it depicting. The dataset is divided into $60000$ training images and $10000$ validation images.This dataset is essentially uniformly distributed, with all classes containing around $9 - 11\%$ of the total dataset.

@@

The Guangzhou Chest X-Ray Pneumonia Dataset consists of $5856$ X-Ray images of children between the ages of one to five. The images were taken from the Woman and Children's Medical Center in Guangzhou, China. The dataset is divided into a training set of $5232$ images and $624$. The dataset is not evenly distributed, having $74.2\%$ instances of pneumonia in the training set and $62.5\%$ instances of pneumonia in the validation set.

>>>>

Convolutional Neural Networks are neural networks which are specialized to work on images, or other types of data where the structure of the data carries meaning. A normal neural network treats each input independently, and inputs are flattened such that any spatial information is lost. CNNs, on the other hand, explicitly take the spatial dimensions into account, by using convolutional kernels and pooling layers as part of their architecture.

@@

A decision tree is an intiutive way of performing both classification and regression tasks, by asking a series of yes or no questions about an instance in the dataset. These questions will split the original dataset, the root node, into child nodes which contain subsets of the dataset, and these nodes will be split again, creating a tree structure which terminates in leaf nodes which imply a choice of class.

For classification problems, the algorithm for creating a decision tree is quite simple. First, consider each feature of the dataset, and attempt to split all instances into two child nodes based on a simple true/false statement about this feature. Having done this for every feature, calculate the degree of homogeneity\footnote{This can described by calculating the Gini-factor, or by calculating the entropy of the set} of the two child nodes created, when considering the actual class of the instances. The feature which created the most homegenous split, is the feature which has the largest effect on the final class, and as such should be the first feature we use to split the dataset. After this, we have two subsets, and we perform the same algorithm on these subsets again. This algorithm terminates when it is not possible to split the data further, because all nodes only contain instances belonging to the same class. In practice, it is advised to stop the algorithm earlier by introducing limits to the depth of the tree, or limits to how few instances can be in one node. Once the algorithm has terminated, we have a series of leaf nodes containing subsets of the data.

@@

To create a random forest, we perform a bootstrap of the original dataset, creating many similar instances with small permutations introduced by resampling with replacement. We now create one decision tree per bootstrap, only using a random subset of the total amount of features every time we perform a split\footnote{The number of features used is often set to $\sqrt{p}$}. The result of this is that we create many slightly different trees. These trees give predictions which are worse than simply creating a single decision tree using all the features and all the data, but when basing our prediction on the majority vote of all the trees, we are able to make predictions which are more generalizable and less prone to overfitting.

@@

The MNIST dataset is a well known multiclass classification problem used to test machine learning algorithms. It consists of $70000$ black and white images of hand drawn digits, each $28$ by $28$ pixels. Each image is labeled with one of ten classes, corresponding to the digit it depicts. The dataset is divided into $60000$ training images and $10000$ validation images. This dataset is essentially uniformly distributed, with all classes containing around $9 - 11\%$ of the total dataset.

@@

The Guangzhou Chest X-Ray Pneumonia Dataset consists of $5856$ X-Ray images of children between the ages of one to five. The images were taken from the Woman and Children's Medical Center in Guangzhou, China. The dataset is divided into a training set of $5232$ images and $624$ validation images. The dataset is not evenly distributed, having $74.2\%$ instances of pneumonia in the training set and $62.5\%$ instances of pneumonia in the validation set.
